---
title: "Exam 2 R Markdown Output"
author: "Grace Mappes"
date: "6/26/2020"
output: word_document
---

```{r setup, include= TRUE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

### call some packages just in case
```{r}
library(dplyr)
library(tidyverse)
library(WDI)
```



# 1. clear environment

```{r}
rm(list=ls(all=TRUE))
```

# 2. call rio and import inequality data set
```{r}
library(rio)
inequality_data <- import("inequality.xlsx")
```

# 3. Type of data set
This is a panel data set because there are multiple
```{r}
# view class of inequality data set
class(inequality_data)
```

# 4. show Denmark and Sweden inequality_gini scores
```{r}
inequality_Denmark <- subset(inequality_data, 
                                 inequality_gini == "Denmark")
inequality_Sweden <- subset(inequality_data, 
                        inequality_gini == "Sweden")
```

# 5. show Brazil inequality_gini scores

```{r}
inequality_Brazil <- subset(inequality_data, inequality_gini == "Brazil")
```

# 6. 
Based on my prior answers, it is better to have a lower inequality_gini score than a higher score.

# 7. View head of data frame
```{r}
head(inequality_data)
```
# 8. Write function accent.remove, apply to Belarus, run head again.
Also changed encoding to UTF-8 in global options before running code.
```{r}
# write function accent.remove
accent.remove <- function(s) {
  old1 <- "ú"
  new1 <- "u"
  s1 <- chartr(old1, new1, s)
}

#apply accent.remove to data set
inequality_data$country = accent.remove(inequality_data$country)

#view head of data set to check for accent removal 
head(inequality_data)
```
# 9. Sort data by lowest inequality_gini score
```{r}
# sort data by lowest inequality_gini score
inequality_data <- order(inequality_data$inequality_gini)

# view head of sorted data set
head(inequality_data)
```

# 10. Mean inequality_gini score
```{r}
mean(inequality_data$inequality_gini, na.rm=TRUE)
```
As seen on the summary table above, the mean inequality_gini score is 36.81.

# 11. Create high_inequality and low_inequality variables with ifelse.
```{r}
#write ifelse
high_inequality <- ifelse(test = inequality_data$inequality_gini > 36.81, yes = 1, no = 0)
low_inequality <- ifelse(test = inequality_data$inequality_gini <= 36.81, yes = 1, no = 0)
```

# 12. Cross tab of high_inequality and low_inequality
```{r}
library(doBy)
summaryBy(high_inequality ~ low_inequality, data = inequality_data, FUN = c(mean,length), 
          na.rm=TRUE, keep.names = TRUE)
```

# 13 For Loop
First, I will make a vector that includes the names of the organizations.
```{r}
x <- c("World Bank", "African Development Bank", "Bill and Melinda Gates Foundation")
```
Next I will write the for loop to print these names.
```{r}
for (x in 1:x) {
  print(paste(x))
}
```

# 14 Find variable from World Development Indicators
I chose the variable rural population because there is a stereotype that rural populations are more impoverished, and I would like to test that assertion. 

# 15. Import rural population variable directly into R
```{r}
# call WDI package to get rural population data
library(WDI)
population_data = WDI(country = "all",
indicator = c("SP.RUR.TOTL"), # indicator from web
start = 2015, end = 2015, extra = FALSE, cache = NULL)
```

# 16. Renaming the variable for clarity
```{r}
# check current variable of name with summary
summary(population_data)
```
```{r}
# load data.table package
library(data.table)
setnames(population_data, "SP.RUR.TOTL", "rural_pop")
```

# 17. Merge the data sets
```{r}
# merge the data sets
merged_df <- dplyr::left_join(inequality_data, population_data)

# check for names that end in .x or .y
names(merged_df)
```

# 18. Remove missing data
```{r}
# remove missing data from inequality_gini
merged_df$inequality_gini <- na.omit(merged_df$inequality_gini)

# remove missing data from new variable, rural_pop
merged_df$rural_pop <- na.omit(merged_df$rural_pop)
```

# 19. Filter data, inequality_gini > 30
```{r}
data_greater_30 <-
  merged_df %>%
  dplyr::filter(inequality_gini > 30)
```

# 20. Count sequence "ai" in country names
```{r}

```

# 21. Sum of inequality_gini in data_greater_30
```{r}
# first subset inequality_gini from data_greater_30
data_greater_30_gini <- subset(data_greater_30, select = "inequality_gini")
# use apply function to create a new variable with sum
data_greater_30_gini_sum <- apply(data_greater_30_gini, 2, sum)
# view sum
View(data_greater_30_gini_sum)
```

# 22. Label merged_df variables
```{r}
# install package labelled
library(labelled)

# label the variables 
var_label(merged_df) <- list(`iso2c` = "Country Code",
`country` = "Country",
`inequality_gini` = "Gini Inequality Score",
`year` = "Year,"
`rural_pop' = "Rural Population")
```

# 23. Save the data as a Stata format
```{r}
# save the dataset in Stata format with the labels
library(rio)
export(merged_df, file = "final_data.dta")
```
